\documentclass[a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[adobe-utopia]{mathdesign}
\usepackage[margin=3cm]{geometry}
\linespread{1.7}

\usepackage[acronym,shortcuts]{glossaries}
\input{acronyms.tex}

\title{Abstract PhD thesis\\ \textsc{Robust Brain-Computer Interfaces}}
\author{Boris Reuderink}

\begin{document}
\maketitle

A \ac{BCI} enables direct communication from the brain to devices,
bypassing the traditional pathway of peripheral nerves and muscles.   
%
Current \acp{BCI} aimed at patients require that the user invests weeks, or even
months, to learn the skill to intentionally modify their brain signals. This can
be reduced to a calibration session of about half an hour per session if
\ac{ML} methods are used. The laborious recalibration is still needed due to
inter-session differences in the statistical properties of the \ac{EEG} signal.
%
Further, the natural variability in spontaneous \ac{EEG} violates basic
assumptions made by the \ac{ML} methods used to train the \ac{BCI} classifier,
and causes the classification accuracy to fluctuate unpredictably. These
fluctuations make the current generation of \acp{BCI} unreliable.
%
In this dissertation, we will investigate the nature of these variations in the
\ac{EEG} distributions, and introduce two new, complementary methods to
overcome these two key issues.
%
To confirm the problem of non-stationary brain signals, we first show that
\acp{BCI} based on commonly used signal features are sensitive to changes in
the mental state of the user.
% 
We proceed by describing a method aimed at removing these changes in signal
feature distributions. We have devised a method that uses a \ac{SOB} to
specifically isolate these relative changes in neuronal firing synchrony. To
the best of our knowledge this is the first \ac{BCI} classifier that works on
out-of-sample subjects without any loss of performance.
%
Still, the assumption made by \ac{ML} methods that the training data consists of
samples that are \ac{IID} is violated, because \ac{EEG} samples nearby in time
are highly correlated. Therefore we derived a generalization of the well-known
\ac{SVM} classifier, that takes the resulting chronological structure of
classification errors into account. Both on artificial data and real \ac{BCI}
data, overfitting is reduced with this \ac{dSVM}, leading to \acp{BCI} with an
increased information throughput.

%With the \ac{SOB} features we have addressed the first key issue (of investment
%of time) for a motor imagery task. The second key issue is partially addressed with the \ac{dSVM}, but raises new questions regarding the
%interpretation of \ac{BCI} performance. 
\end{document}
