\chapter{Cross-subject generalization}
%\chapter{Removing inter-session variability}
\label{chap:sob}

% motivation to look at changing feature dists
\lettrine{I}{n} Chapter~\ref{chap:loc} we demonstrated that \acp{BCI} based on
oscillatory (\ac{ERD}) features are sensitive to changes in the mental
state --- changes that are unrelated to the mental task used to control the
\ac{BCI}. While the \ac{ERP} based classifiers were proven to result in higher
and more reliable performance than their \ac{ERD} counterparts, here we focus
on improving the \ac{ERD} based classifiers that exhibit the problematic drifts
in performance well known in the \ac{BCI} field\footnote{This chapter is based
on \cite{reuderink2011sib}, accepted for oral presentation at EMBC'2011.}. The
reason we focus on a \ac{ERD} based \ac{BCI} is that it is well studied, and
known to work when the user only imagines movement, and does not actually
perform movement as was the case in Chapter~\ref{chap:loc}; for \ac{LRP}
classification we are not aware of \ac{BCI} studies with fully imagined
movement. Furthermore, improvement obtained with movement-related \ac{ERD}
might prove to be beneficial for other oscillatory brain signatures, such as
those related to covert attention \citep[e.g.][]{vangerven2009amp}.

% drifts in joint feature distributions
Changes in \ac{BCI} performance might be caused by changes
in the generating processes (e.g. due to changes in attention, or different
processes for motor control), or more generally, by changes in the joint
distribution of task-relevant and task-irrelevant \ac{EEG} features. This means
that not only changes in the task-relevant neural activity, but also unrelated
changes (e.g. \ac{LOC} in the previous chapter) might affect the \ac{BCI}.
% introduce generalization
Changes or drifts in the feature distributions pose a problem for automatic
discrimination, since the ability to generalize to new data of a static
\ac{BCI} classifier depends on the appropriateness of the learned decision
function for the online data. When there is a change in the feature
distribution compared to the calibration data (covariate shift), it is likely
that the quality of the predictions will degrade; the \ac{BCI} classifier does not
generalize well to this data.
% 
This difficulty of generalization holds for application of a trained
\ac{BCI} to future data within the same session, to new sessions of the same
subject, and to other subjects, in increasing order of difficulty.

% Why relevant
Being able to successfully apply a \ac{BCI} to new sessions and subjects is
relevant for various reasons. Practically, a \ac{BCI} that does not require a
calibration session before each use is much more user friendly. Secondly, if a
\ac{BCI} generalizes to new users this indicates that the mechanism used to
classify is a generic one. This aids interpretation of the neural signatures
used for control. And finally, if a \ac{BCI} is known to operate robustly
during changing contexts (i.e. mental states, recording environment), this
means its operation is not depending on these contexts.

% Adaptation in feat-space vs in classifier
To achieve generalization under covariate-shift, either the classifier can be
adap\-ted to the changing feature distributions \cite{vidaurre2011tua,
shenoy2006tac}, or the spatial filters themselves can be adapted
\cite{bunau2009ssa, meinecke2009lis, tomioka2006asf, hill2006tdd, satti2010csm,
blankertz2007ics, shenoy2006tac}. The advantage of adapting the spatial filters
is that a wide variety (e.g. disconnected sensors) of distribution changes can
be tracked when they are estimated correctly. While adaptation of the
classifier cannot handle all these changes, due to its simplicity it is easier
to estimate, and can therefore be more robust.
%
A further distinction can be made between methods that perform adaptation only
during the (off-line) training phase \citep[e.g.][]{bunau2009ssa,
meinecke2009lis, blankertz2007ics}, and methods that perform online adaptation
\cite{vidaurre2011tua, tomioka2006asf, hill2006tdd, satti2010csm,
shenoy2006tac}. The off-line adaptation methods implicitly assume that there is
a spatial subspace that is stationary, and retains most of the discriminatory
information. It is unclear whether such a subspace does indeed exist.
%
Therefore, we took the most flexible approach, that combines online adaptation
with adaptation in feature space (i.e. adaptation of the spatial filters).

% Scope and summary of our method
In this chapter, we will present a \ac{SOB} procedure that reduces long-term,
non task-related variations. This procedure enables the creation of a \ac{BCI}
that can be applied to new subjects \emph{without a calibration session}.
%
Since non-stationary distributions of specifically \ac{ERD} based features have
been reported in \cite{shenoy2006tac, blankertz2007ics, jatzev2008ecn} and are
supported by the results presented in Chapter~\ref{chap:loc}, we evaluate our
method with \ac{ERD} based features --- although it is in no way limited to
this type of feature extraction.

In the next sections, we will describe prior work on subject-independent
classification, outline our baselining method and describe an off-line
experiment used to assess the performance on unseen subjects. Then we will
describe and discuss the results, and end with conclusions and remarks for
future work.
\clearpage % against underfull \vbox

\input{tex/sec_prevwork.tex}
\input{tex/sec_methods.tex}
\input{tex/sec_results.tex}
\input{tex/sec_discussion.tex}
\input{tex/sec_conclusion.tex}
